A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-05-11 paper/paper.Rmd
section 2 done

Our aim is to refresh the wages textbook data and append it with data from 1994 through to the latest data reported in 2018, a purpose consistent with @grimshaw's statistics education goal of embracing authentic data experiences. Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by @SingerJudithD2003Alda. We should also note that race is a variable in the original data set, and for compatibility, it is also provided with the refreshed data for the *purposes of studying racism, not race* [@racismnotrace]. There are a number of data sets provided by @SingerJudithD2003Alda, and we focus only on this one because it has captivated our attention for a number of years. We use it in our own teaching of longitudinal data analysis and would very much like to use data since 1994. <!--However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.-->

@@

1. The wages data of the whole NLSY79 cohort, including females.
2. A separate table of the demographic data of the whole NLSY79 cohort.
3. The subset of wages data in (1), which is the wages of high school dropouts' as a refreshed version of @SingerJudithD2003Alda's data.

@@

In the context of official statistics, @LooMarkvander2018Sdcw describe the "statistical value chain," which includes various production stages of the data cleaning process as raw data (data in the initial that it arrives), input data (data organized with correct type and identified variables), and valid data (data that has been cleaned and more accurately represents the intent of variables). What we have colorfully named wild data can be considered to be raw data, and valid data could be considered to be textbook data in the above statistical value chain. In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidy the raw data into input data, specifically for the demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida) and Section \@ref(censor).

>>>>

Our aim is to refresh the wages textbook data and append it with data from 1994 through to the latest data reported in 2018, a purpose consistent with @grimshaw's statistics education goal of embracing authentic data experiences. Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by @SingerJudithD2003Alda. We should also note that race is a variable in the original data set, and for compatibility, it is also provided with the refreshed data for the *purposes of studying racism, not race* [@racismnotrace]. There are a number of data sets provided by @SingerJudithD2003Alda, but we focus only on wages data because it has captivated our attention for a number of years in our own teaching of longitudinal data analysis. <!--However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.-->

@@

1. The wages data of the whole NLSY79 cohort, including females,
2. A separate table of the demographic data of the whole NLSY79 cohort, and 
3. The subset of wages data in (1), which is the wages of high school dropouts' as a refreshed version of @SingerJudithD2003Alda's data.

@@

In the context of official statistics, @LooMarkvander2018Sdcw describe the "statistical value chain," which includes various production stages of the data cleaning process as raw data (initial data as delivered originally), input data (data organized with correct type and identified variables), and valid data (data that has been cleaned and more accurately represents the intent of variables). What we have colorfully named wild data can be considered to be raw data, and valid data could be considered to be textbook data in the above statistical value chain. In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidy the raw data into input data, specifically for the demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida) and Section \@ref(censor).
